Simulation Learning in Computer Hex

### 2011-10-17 thesis/thesis.tex
Changed from png graphics to havannah.sty board defs.

\section{Background: The Game of Hex}
Hex is a two player, full-information, abstract strategy game played on an $N$x$N$ grid of hexagons. The players are denoted by colors here \black{} and \white. The object of the game is to form an unbroken chain from one side of the board to the other. \black{} attempts to connect top to bottom, \white{} left to right. Figure \ref{fig:13x13inprogress} shows an example game in progress on a 13x13 board. Figure \ref{fig:13x13finished} shows the final position, \black{} having just played the winning move, with an unbroken chain connecting top to bottom. Hex has properties that make it attractive for AI research. Once placed, a piece never leaves the board. The only criterion for a legal move is that the hexagon be unoccupied (empty). Gale '79 proved the game cannot end in a draw, simplifying minimax methods. Nash, one of the inventors of the game, used a strategy stealing argument to prove the existence of a winning strategy for the first player, but the proof is by reduction and offers no insight into the strategy itself. Solutions for all opening positions for 7x7 and 8x8 boards have been found\cite{henderson2009solving}. A good portion, but not all, 9x9 openings have also been solved. Nash proposed 11x11 as the ideal board size, but 13x13 has been more common in competitions, especially as the ability of computer players improves.

@@

\begin{figure}
	\begin{center}
	\includegraphics[width=200px]{graphics/13x13-in-progress.png}
	\end{center}
	\caption{13x13 Hex Game, \white{} to play.}
	\label{fig:13x13inprogress}
\end{figure}
\begin{figure}
	\begin{center}
	\includegraphics[width=200px]{graphics/13x13-finished.png}
	\end{center}
	\caption{Finished 13x13 Hex Game, \black{}'s play at \loc{b6} gives him an unbroken path from top to bottom.}
	\label{fig:13x13finished}
\end{figure}

>>>>

\section{Background: The Game of Hex}
Hex is a two player, full-information, abstract strategy game played on an $N$x$N$ grid of hexagons. The players are denoted by colors here \black{} and \white. The object of the game is to form an unbroken chain from one side of the board to the other. Figure \ref{fig:13x13inprogress} shows an example game in progress on a 13x13 board. Figure \ref{fig:13x13finished} shows the final position, \black{} having just played the winning move, with an unbroken chain connecting both his sides. Hex has properties that make it attractive for AI research. Once placed, a piece never leaves the board. The only criterion for a legal move is that the hexagon be unoccupied (empty). Gale '79 proved the game cannot end in a draw, simplifying minimax methods. Nash, one of the inventors of the game, used a strategy stealing argument to prove the existence of a winning strategy for the first player, but the proof is by reduction and offers no insight into the strategy itself. Solutions for all opening positions for 7x7 and 8x8 boards have been found\cite{henderson2009solving}. A good portion, but not all, 9x9 openings have also been solved. Nash proposed 11x11 as the ideal board size, but 13x13 has been more common in competitions, especially as the ability of computer players improves.

@@

\begin{figure}
	\begin{HexBoard}[board size=13]
		\HGame{c4,f6,e11,i8,h8,i1,i6,d5,i3,
					 h9,h5,h4,i4,g9,f9,g8,f7,f8,e8}
	\end{HexBoard}
	\caption{13x13 Hex Game, \white{} to play.}
	\label{fig:13x13inprogress}
\end{figure}

\begin{figure}
	\begin{HexBoard}[board size=13]
		\HGame{c4,f6,e11,i8,h8,i1,i6,d5,i3,
				   h9,h5,h4,i4,g9,f9,g8,f7,f8,
				   e8,e9,c10,k8,c9,g6,g7,c11,
				   d10,h6,h7,i5,j5,j4,k4,k3,
				   m2,l3,m3,l4,c6,e11,l5,m4,
				   d11,b7,e6,d12,c12,c7,e7,f5,
				   e5,d9,d8,f4,e4,f3,a2,e3,d6,
				   c3,d3,d4,b5,e2,d2,d1,e1,b13,
				   c13,c5,b6}
	\end{HexBoard}
	\caption{Finished 13x13 Hex Game, \black{}'s play at \loc{b6} gives him an unbroken path from top to bottom.}
	\label{fig:13x13finished}
\end{figure}
